Security Properties for Stack Safety

### 2021-05-15 short/main.tex
Cleaned up some spot, tried to make para. 4 better

Our threat model is very strong, allowing the attacker (i.e., the callee) to
execute arbitrary code, including code that might attempt to smash
the stack to disrupt the program's control flow.
Our stack safety properties demand that, even in the presence of such
attackers, confidentiality and integrity must
still apply to data in suspended stack frames until the callee returns.
% all of the data they would under more idealized conditions.
%This formulation not only captures the intuition
%that the callee cannot directly access the caller's state; it also gives a novel
%interpretation of integrity and confidentiality in the presence of control-flow
%attacks.
% \bcp{I tried rewriting the first part of the paragraph, and I like it
%   better, but now there's a big gap between that and what comes next.  I
%   wonder whether we need to turn the next bit into a separate discussion of
%   ``what it means to return''...  In any case, I felt like the previous
%   version of the whole paragraph didn't quite nail what we needed to say
%   about all this.}
% \sna{Better transition.}
The question then arises: when, exactly, are suspended frames
reactivated---or equivalently,
what constitutes a return in this setting? It is not obvious, since our
setting is a concrete, low-level ISA where the same opcode might sometimes
be used to return from a call and sometimes for other purposes.
Our solution is to associate with each call a corresponding predicate on states,
its ``return target,''
which defines what it means for that call to truly return: the program counter
must be at the next instruction, and the stack pointer restored to its original
value. A callee may execute a return instruction,
but if it does not reach a return target we simply do not recognize it as a
return.
So the callee can be ``active'', and the caller protected, even if the callee is
executing instructions outside of its normal footprint.  \bcp{I tried
  another rewrite of the end of the paragraph.  I think it's getting there,
  but it's probably now a bit too long as well as too detailed.}

%Our attacker model is strong, allowing the attacker to execute arbitrary code, so
%the callee has no reason to execute the caller's instructions -- any instruction it could
%jump to, it could already execute. The only reason to perform a ``fake'' return in this setting
%would be to access the caller's data, which our properties will correctly identify as a violation.

These properties can optionally be extended with a notion of {\em well-bracketed control flow}
as in \citet{SkorstengaardSTK}---a global requirement that callees
must always return
to their immediate caller, if they return at all. This property is orthogonal to our
data-protection properties. It remains meaningful in our strong attacker model, because
it supports reasoning about the ordering of valid returns.

To demonstrate the utility of our properties, we use them
to evaluate an existing mechanism, the
{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}, re-implemented
in the Coq proof assistant on top of a RISC-V ISA model. We
use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for Coq, to generate random programs and check
that Roessler and Dehon's micro-policies correctly abort programs that
would violate stack safety.
Our testing regime is also able to detect incorrect variants of the
enforcement mechanisms---both variants that we accidentally created
during our re-implementation of the micro-policy and ones that we
intentionally crafted to be broken in order to increase our confidence
in testing and the enforcement mechanism itself. For these broken
mechanisms, our tester
generates counterexamples that violate our properties but escape the
enforcement.  This increases our confidence in enforcement
mechanisms for which the tester is unable to find  counterexamples.

@@

In ongoing work, we are investigating whether stack protection schemes
based on the Cheri capability system~\citep{Woodruff+14,Chisnall+15}
can enforce our properties.  In particular, we are adapting
the ``uninitialized capability model'' of \citet{Georges+21} to work with
our testing framework in order to validate it against our properties. Out of several Cheri-based
stack protection schemes, this one's treatment of unitialized frames seems most compatible
with our model of confidentiality, without the need for expensive stack clearing
between calls. We expect that Cheri-based techniques can support our properties if we
assume extra cooperation from callers, namely that they do not leak their capabilities
to the attacker. %There is no equivalent dynamic requirement in a tag-based system because
%the policy can dynamically enforce cooperation.

>>>>

Our threat model is very strong, allowing the attacker (i.e., the callee) to
execute arbitrary code, including code that might attempt to smash
the stack to disrupt the program's control flow.
Our stack safety properties demand that, even in the presence of such
attackers, confidentiality and integrity must
still apply to data in suspended stack frames until the callee returns.
The question then arises:
what constitutes a return in this setting? It is not obvious, since our
setting is a concrete, low-level ISA where the same opcode might sometimes
be used to return from a call and sometimes for other purposes.
Our solution is to associate with each call a corresponding predicate on states,
its ``return target,''
which holds on the states that would follow a well-behaved return.
Whether or not a callee executes a return instruction,
we only recognize it returning only when it reaches a return target.
So the caller is protected until then, even if the callee is
executing instructions outside of its normal footprint.

These properties can optionally be extended with a notion of {\em well-bracketed control flow}
as in \citet{SkorstengaardSTK}---a global requirement that callees
must always return
to their immediate caller, if they return at all. This property is orthogonal to our
data-protection properties. It remains meaningful in our strong attacker model,
to support reasoning about the ordering of valid returns.

To demonstrate the utility of our properties, we use them
to evaluate an existing mechanism, the
{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}, re-implemented
in the Coq proof assistant on top of a RISC-V ISA model. We
use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for Coq, to generate random programs and check
that Roessler and Dehon's micro-policies correctly abort programs that
would violate stack safety.
Our testing regime also detects incorrect variants of the
enforcement mechanisms---both errors
during our re-implementation of the micro-policy and
intentionally crafted test, meant to be broken in order to increase our confidence
in testing and the enforcement mechanism itself. For these broken
mechanisms, our tester
generates counterexamples that violate our properties but escape the
enforcement.

@@

In ongoing work, we investigate whether stack protection schemes
based on the Cheri capability system~\citep{Woodruff+14,Chisnall+15}
can enforce our properties.  In particular, we aim to validate
the ``uninitialized capability model'' of \citet{Georges+21} using
our testing framework. Out of several Cheri-based
stack protection schemes, their treatment of unitialized frames seems most compatible
with our model of confidentiality, without the need for expensive stack clearing
between calls. We expect that Cheri-based techniques can support our properties if we
assume extra cooperation from callers, namely that they do not leak their capabilities
to the attacker. %There is no equivalent dynamic requirement in a tag-based system because
%the policy can dynamically enforce cooperation.
